---
title: "Cluster analysis"
author: "Mattias Bengtsson"
date: "2024-04-25"
output:
  pdf_document:
    toc: no
  bookdown::pdf_document2:
    toc: no
    fig_cap: yes
    keep_tex: yes
    fontsize: 12pt
    bibliography: ref.bib
    csl: "https://www.zotero.org/styles/chicago-fullnote-bibliography"
    suppress-bibliography: yes
  word_document:
    toc: no
header-includes:
- \usepackage{float}
- \floatplacement{table}{H}
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = FALSE
)
```


# Setup

```{r}

library(tidyverse)
library(FactoMineR)
library(factoextra)
library(explor)
library(performanceEstimation)
library(nnet)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(modelsummary)
library(kableExtra)
library(flextable)
library(cluster)
library(tinytable)
```


```{r}
setwd("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/Clusteranalysis")#liu dator
```
```{r}
#setwd("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/Clusteranalysis") 
```


```{r}
imputed_data_0121 <- read.csv("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/Clusteranalysis/combined_imputed_data0121_incomeandcult.csv")

#imputed_data_0121 <- read.csv("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/Clusteranalysis/combined_imputed_data0121_incomeandcult.csv")



```


```{r}
data <- mutate_at(imputed_data_0121,
                           vars(-year, -age, -idnr), 
                           as.factor)
```


```{r}
# Excluding parties that are not in the national parliament
data <- data %>% 
  filter(cb10 != "11" & cb10 != "12")

# Updating levels
data$cb10 <- droplevels(data$cb10)

nlevels(data$cb10)
```


```{r}
# Reordering variables so that they following the logic of other variables, where 1 is the lowest value. 
data <- data |> 
  mutate(p1 = case_when(
    jc10a == "1" ~ "6",
    jc10a == "2" ~ "5",
    jc10a == "3" ~ "4",
    jc10a == "4"~ "3",
    jc10a == "5" ~ "2",
    jc10a == "6" ~ "1"
  ))

data <- data |> 
  mutate(euroesec_2 = case_when(
    euroesec == "1" ~ "9",
    euroesec == "2" ~ "8",
    euroesec == "3" ~ "7",
    euroesec == "4"~ "6",
    euroesec == "5" ~ "5",
    euroesec == "6" ~ "4",
    euroesec == "7" ~ "3",
    euroesec == "8" ~ "2",
    euroesec == "9" ~ "1"
  ))

# Making sex so that 1 = man, 0 = woman
data <- data %>% 
  mutate(sex_binary = case_when(
    sex == "1" ~ 0,
    sex == "2" ~ 1
  ))

# Renaming variables so that they are understandable
data <- data %>%  
  mutate(praying = bc20,
         Cinema = la10,
         Rock_pop = lb10b,
         Theatre = lb10a,
         Classical_opera = lb10e,
         Library = eg10,
         Pub_restaurant = ma50c,
         Drawn_painted = le10c,
         Friends = ma10a,
         Church = bc10,
         snus = ma40h,
         gamble_lottery = mc10c,
         Generations = genzukin,
         Subjective_class_today = subclg, 
         Income = hinc5rel,
         Educationlevel = edu3
         )
```

# Clustering 2019

```{r}
# Filtering out data from 2019, and choosing variables. 
Clusteringdata_2019 <- data %>% 
    filter(year == 2019) %>% 
  select(idnr,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1
)

# Here i add the idnr to rownames,so that we can merge the values for the ID with their values on variables that are not included in the cluster analysis. 
row.names(Clusteringdata_2019) <- Clusteringdata_2019$idnr

Clusteringdata_2019 <- Clusteringdata_2019 %>% 
  select(-idnr)
```

```{r}
# Making a distance matrix with default settings. 
dist_matrix_2019 <- dist(Clusteringdata_2019)

```

```{r}
# Performing hierarchical clustering with ward.D2. 
set.seed(123)
cluster_result_2019 <- hclust(dist_matrix_2019, method = "ward.D2")

# Visualize dendrogram
dendogram2019 <- plot(cluster_result_2019, main = "Hierarchical Clustering Dendrogram 2019",
     xlab = "Observations", sub = NULL)


```
```{r}
# Saving dendogram
png("Dendogram2019.png", width = 2000, height = 1400, res = 300)
plot(cluster_result_2019, main = "Hierarchical Clustering Dendrogram 2019",
     xlab = "Observations", sub = NULL)
dev.off()
```



# Clustering  2014


```{r}
# Filtering out data from 2014, and choosing variables. 
Clusteringdata_2014 <- data %>% 
    filter(year == 2014) %>% 
  select(idnr,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1
)

# Here i add the idnr to rownames,so that we can merge the values for the ID with their values on variables that are not included in the cluster analysis.
row.names(Clusteringdata_2014) <- Clusteringdata_2014$idnr

Clusteringdata_2014 <- Clusteringdata_2014 %>% 
  select(-idnr)
```

```{r}
# Making a distance matrix with default settings. 
dist_matrix_2014 <- dist(Clusteringdata_2014)

```

```{r}
# Performing hierarchical clustering with ward.D2. 
set.seed(123)

cluster_result_2014 <- hclust(dist_matrix_2014, method = "ward.D2")

# Visualize dendrogram
dendogram2014 <- plot(cluster_result_2014, main = "Hierarchical Clustering Dendrogram 2014",
     xlab = "Observations", sub = NULL)

```
```{r}
png("Dendogram2014.png", width = 2000, height = 1400, res = 300)
plot(cluster_result_2014, main = "Hierarchical Clustering Dendrogram 2014",
     xlab = "Observations", sub = NULL)
dev.off()
```



# Clustering 2008


```{r}
# Filtering out data from 2008, and choosing variables. 
Clusteringdata_2008 <- data %>% 
    filter(year == 2008) %>% 
  select(idnr,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1
)

# Here i add the idnr to rownames,so that we can merge the values for the ID with their values on variables that are not included in the cluster analysis.
row.names(Clusteringdata_2008) <- Clusteringdata_2008$idnr

Clusteringdata_2008 <- Clusteringdata_2008 %>% 
  select(-idnr)
```

```{r}
# Making a distance matrix with default settings. 
dist_matrix_2008 <- dist(Clusteringdata_2008)

```

```{r}
# Performing hierarchical clustering with ward.D2. 
set.seed(123)

cluster_result_2008 <- hclust(dist_matrix_2008, method = "ward.D2")

# Visualize dendrogram
dendogram2008 <- plot(cluster_result_2008, main = "Hierarchical Clustering Dendrogram 2008",
     xlab = "Observations", sub = NULL)

```
```{r}
png("Dendogram2008.png", width = 2000, height = 1400, res = 300)
plot(cluster_result_2008, main = "Hierarchical Clustering Dendrogram 2008",
     xlab = "Observations", sub = NULL)

dev.off()
```



# Descriptive statistics of clusters K = 8


## 2008


```{r}
# Cut dendrogram into 8 clusters representing the number of parties. 
clusters_2008_k8 <- cutree(cluster_result_2008, k = 8)

clusters_2008_k8 <- as.data.frame(clusters_2008_k8)

# Setting rownnames as idnr, thereby we can merge together with the original data on Idnr
clusters_2008_k8$idnr <- row.names(clusters_2008_k8)

row.names(clusters_2008_k8) <- NULL
```

```{r}
# Taking the original data, filtered on year 2008
data2008_k8 <- data %>% 
  filter(year == 2008)
```

```{r}
# Merging cluster assignment with original data, by idnr
data2008_k8 <- merge(data2008_k8, clusters_2008_k8, by = "idnr")
```


#### Party proportions
```{r}
# Here I create separate vectors for each cluster to be able to calculate proportions. 
Clusteringdata_2008k8_cluster1 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "1")

Clusteringdata_2008k8_cluster2 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "2")

Clusteringdata_2008k8_cluster3 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "3")

Clusteringdata_2008k8_cluster4 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "4")

Clusteringdata_2008k8_cluster5 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "5")

Clusteringdata_2008k8_cluster6 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "6")

Clusteringdata_2008k8_cluster7 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "7")

Clusteringdata_2008k8_cluster8 <- data2008_k8 %>% 
  filter(clusters_2008_k8 == "8")
```

```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc1_k8 <- prop.table(table(Clusteringdata_2008k8_cluster1$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc1_k8 <- as.data.frame(category_proportionsc1_k8)

names(category_proportions_dfc1_k8) <- c("Party", "Proportion_C1")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 2
category_proportionsc2_k8 <- prop.table(table(Clusteringdata_2008k8_cluster2$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc2_k8 <- as.data.frame(category_proportionsc2_k8)

names(category_proportions_dfc2_k8) <- c("Party", "Proportion_C2")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 3
category_proportionsc3_k8 <- prop.table(table(Clusteringdata_2008k8_cluster3$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc3_k8 <- as.data.frame(category_proportionsc3_k8)

names(category_proportions_dfc3_k8) <- c("Party", "Proportion_C3")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 4
category_proportionsc4_k8 <- prop.table(table(Clusteringdata_2008k8_cluster4$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc4_k8 <- as.data.frame(category_proportionsc4_k8)

names(category_proportions_dfc4_k8) <- c("Party", "Proportion_C4")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 5
category_proportionsc5_k8 <- prop.table(table(Clusteringdata_2008k8_cluster5$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc5_k8 <- as.data.frame(category_proportionsc5_k8)

names(category_proportions_dfc5_k8) <- c("Party", "Proportion_C5")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc6_k8 <- prop.table(table(Clusteringdata_2008k8_cluster6$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc6_k8 <- as.data.frame(category_proportionsc6_k8)

names(category_proportions_dfc6_k8) <- c("Party", "Proportion_C6")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc7_k8 <- prop.table(table(Clusteringdata_2008k8_cluster7$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc7_k8 <- as.data.frame(category_proportionsc7_k8)

names(category_proportions_dfc7_k8) <- c("Party", "Proportion_C7")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc8_k8 <- prop.table(table(Clusteringdata_2008k8_cluster8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc8_k8 <- as.data.frame(category_proportionsc8_k8)

names(category_proportions_dfc8_k8) <- c("Party", "Proportion_C8")

```


```{r}
# Calculate proportions of categories from cb10 for all

category_proportionsk8 <- prop.table(table(data2008_k8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfk8 <- as.data.frame(category_proportionsk8)

names(category_proportions_dfk8) <- c("Party", "Proportion_all")
```

```{r}
# Merging them all to together
merged_2008_k8 <- merge(category_proportions_dfc1_k8, category_proportions_dfc2_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc3_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc4_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc5_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc6_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc7_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfc8_k8, by = "Party")

merged_2008_k8 <- merge(merged_2008_k8 , category_proportions_dfk8, by = "Party")

```

```{r}
# Renaming to ful party name
merged_2008_k8 <- merged_2008_k8 %>% 
  mutate(Party = case_when(
    Party == 1 ~ "Left Party",
    Party == 2 ~ "Social Democrats",
    Party == 3 ~ "Centre Party",
    Party == 4 ~ "Liberals",
    Party == 5 ~ "Moderate Party",
    Party == 6 ~ "Christian Democrats",
    Party == 7 ~ "Green Party",
    Party == 10 ~ "Sweden Democrats"
  ))

# Renaming for clarity
merged_2008_k8 <- merged_2008_k8 %>% 
  rename("Cluster 1" = Proportion_C1,
         "Cluster 2 " = Proportion_C2,
          "Cluster 3" = Proportion_C3,
            "Cluster 4" = Proportion_C4,
         "Cluster 5" = Proportion_C5,
         "Cluster 6" = Proportion_C6,
         "Cluster 7" = Proportion_C7,
         "Cluster 8" = Proportion_C8,
         "All" = Proportion_all)

# Taking the proportion *100 to get percentages. 
merged_2008_k8 <- merged_2008_k8 %>%
  mutate_at(vars(starts_with("Cluster")), funs(. * 100))

merged_2008_k8 <- merged_2008_k8 %>%
  mutate_at(vars(starts_with("All")), funs(. * 100))

```

```{r}
# Printing the table for proportions of respondent in each cluster that prefers each party. Output in latex used for Overleaf. 
table_proportions_cluster_2008k8 <- datasummary_df(merged_2008_k8, fmt = 0,
               title = "Preferred party of respondents in each cluster, for the year 2008",
               note = "Data from SOM",
               output = "latex")
```


```{r echo=FALSE, include = TRUE}
table_proportions_cluster_2008k8
```


#### Descriptives of lifestyle and class variables

Here I create the descriptive tables which show the mean values of different lifestyle activities, as well as for class variables, for each cluster.

```{r}
#  Choosing all variables, now adding class variables and demographics. 
data2008_tabledata_k8 <- data2008_k8 %>% 
  select(clusters_2008_k8,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1,
         euroesec_2,
         Income,
         Educationlevel,
         Subjective_class_today,
         age,
         sex_binary)


```

```{r}
# Converting to dataframe
data2008_tabledata_k8 <- as.data.frame(data2008_tabledata_k8)

# Making variables numeric to calculate means
data2008_tabledata_means_k8 <- data2008_tabledata_k8 %>%
  mutate_all(as.numeric)

# Making cluster names character
data2008_tabledata_means_k8 <- data2008_tabledata_means_k8 %>%
  mutate(clusters_2008_k8 = as.character(clusters_2008_k8))

# Renaming cluster names
data2008_tabledata_means_k8 <- data2008_tabledata_means_k8 %>% 
  mutate(cluster = case_when(
    clusters_2008_k8 == "1" ~ "C 1",
    clusters_2008_k8 == "3" ~ "C 2",
    clusters_2008_k8 == "2" ~ "C 3",
    clusters_2008_k8 == "6" ~ "C 4",
    clusters_2008_k8 == "5" ~ "C 5",
    clusters_2008_k8 == "4" ~ "C 6",
    clusters_2008_k8 == "3" ~ "C 2",
    clusters_2008_k8 == "7" ~ "C 7",
    clusters_2008_k8 == "8" ~ "C 8"))

```

```{r echo=FALSE, include = TRUE}
# Table with descriptive statistics of the clusters using Mean. Output in latex to use in Overleaf. 

datasummary(praying+
         Cinema+
         Rock_pop+
         Theatre+
         Classical_opera+
         Library+
         Pub_restaurant+
         Drawn_painted+
         Friends+
         Church+
         snus+
         gamble_lottery+
         p1+
         euroesec_2+
         Income+
         Educationlevel+
         Subjective_class_today+
         age+
         sex_binary ~ cluster * (Mean) + Mean,
            data = data2008_tabledata_means_k8,
         title = "2008",
         output = "latex")



```
### Descriptive table
```{r}
# Descriptives showing the percentages of observations that reside in each cluster. Output in latex to use in Overleaf
datasummary(
         sex_binary ~ cluster * (Percent()) + N, #I use sex_binary as a placeholder to create a table. 
            data = data2008_tabledata_means_k8,
         title = "2008",
         output = "latex")

```

## 2014


```{r}
# Cut dendrogram into 8 clusters representing the number of parties.
clusters_2014_k8 <- cutree(cluster_result_2014, k = 8)

clusters_2014_k8 <- as.data.frame(clusters_2014_k8)

# Setting rownames as idnr, thereby we can merge together with the original data on idnr.

clusters_2014_k8$idnr <- row.names(clusters_2014_k8)

row.names(clusters_2014_k8) <- NULL
```

```{r}
# Taking the original data, filtered on year 2014.
data2014_k8 <- data %>% 
  filter(year == 2014)
```

```{r}
# Merging cluster assignment with original data by idnr.
data2014_k8 <- merge(data2014_k8, clusters_2014_k8, by = "idnr")
```


#### Party proportions
```{r}
# Here I create separate vectors for each cluster to be able to calculate proportions.

Clusteringdata_2014k8_cluster1 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "1")

Clusteringdata_2014k8_cluster2 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "2")

Clusteringdata_2014k8_cluster3 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "3")

Clusteringdata_2014k8_cluster4 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "4")

Clusteringdata_2014k8_cluster5 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "5")

Clusteringdata_2014k8_cluster6 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "6")

Clusteringdata_2014k8_cluster7 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "7")

Clusteringdata_2014k8_cluster8 <- data2014_k8 %>% 
  filter(clusters_2014_k8 == "8")
```

```{r}
# Calculating proportions of categories from cb10 for cluster 1.

category_proportionsc1_k8 <- prop.table(table(Clusteringdata_2014k8_cluster1$cb10))

# Converting the result to a data frame for plotting.
category_proportions_dfc1_k8 <- as.data.frame(category_proportionsc1_k8)

names(category_proportions_dfc1_k8) <- c("Party", "Proportion_C1")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 2
category_proportionsc2_k8 <- prop.table(table(Clusteringdata_2014k8_cluster2$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc2_k8 <- as.data.frame(category_proportionsc2_k8)

names(category_proportions_dfc2_k8) <- c("Party", "Proportion_C2")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 3
category_proportionsc3_k8 <- prop.table(table(Clusteringdata_2014k8_cluster3$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc3_k8 <- as.data.frame(category_proportionsc3_k8)

names(category_proportions_dfc3_k8) <- c("Party", "Proportion_C3")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 4
category_proportionsc4_k8 <- prop.table(table(Clusteringdata_2014k8_cluster4$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc4_k8 <- as.data.frame(category_proportionsc4_k8)

names(category_proportions_dfc4_k8) <- c("Party", "Proportion_C4")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 5
category_proportionsc5_k8 <- prop.table(table(Clusteringdata_2014k8_cluster5$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc5_k8 <- as.data.frame(category_proportionsc5_k8)

names(category_proportions_dfc5_k8) <- c("Party", "Proportion_C5")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc6_k8 <- prop.table(table(Clusteringdata_2014k8_cluster6$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc6_k8 <- as.data.frame(category_proportionsc6_k8)

names(category_proportions_dfc6_k8) <- c("Party", "Proportion_C6")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc7_k8 <- prop.table(table(Clusteringdata_2014k8_cluster7$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc7_k8 <- as.data.frame(category_proportionsc7_k8)

names(category_proportions_dfc7_k8) <- c("Party", "Proportion_C7")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc8_k8 <- prop.table(table(Clusteringdata_2014k8_cluster8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc8_k8 <- as.data.frame(category_proportionsc8_k8)

names(category_proportions_dfc8_k8) <- c("Party", "Proportion_C8")

```


```{r}
# Calculate proportions of categories from cb10 for all

category_proportionsk8 <- prop.table(table(data2014_k8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfk8 <- as.data.frame(category_proportionsk8)

names(category_proportions_dfk8) <- c("Party", "Proportion_all")
```

```{r}
# Merging all proportion data frames together.
merged_2014_k8 <- merge(category_proportions_dfc1_k8, category_proportions_dfc2_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc3_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc4_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc5_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc6_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc7_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfc8_k8, by = "Party")

merged_2014_k8 <- merge(merged_2014_k8 , category_proportions_dfk8, by = "Party")

```

```{r}
# Renaming party codes to full party names.
merged_2014_k8 <- merged_2014_k8 %>% 
  mutate(Party = case_when(
    Party == 1 ~ "Left Party",
    Party == 2 ~ "Social Democrats",
    Party == 3 ~ "Centre Party",
    Party == 4 ~ "Liberals",
    Party == 5 ~ "Moderate Party",
    Party == 6 ~ "Christian Democrats",
    Party == 7 ~ "Green Party",
    Party == 10 ~ "Sweden Democrats"
  ))

# Renaming for clarity.
merged_2014_k8 <- merged_2014_k8 %>% 
  rename("C 1 " = Proportion_C1,
         "C 2 " = Proportion_C2,
         "C 3 " = Proportion_C3,
         "C 4 " = Proportion_C4,
         "C 5 " = Proportion_C5,
         "C 6 " = Proportion_C6,
         "C 7 " = Proportion_C7,
         "C 8 " = Proportion_C8,
         "All " = Proportion_all)

# Taking the proportion * 100 to get percentages.
merged_2014_k8 <- merged_2014_k8 %>%
  mutate_at(vars(starts_with("C")), funs(. * 100))

merged_2014_k8 <- merged_2014_k8 %>%
  mutate_at(vars(starts_with("All")), funs(. * 100))

```


```{r}
# Printing the table for proportions of respondents in each cluster that prefer each party. Output in latex for Overleaf.

table_proportions_cluster_2014k8 <- datasummary_df(merged_2014_k8, fmt = 0,
               title = "Preferred party of respondents in each cluster, k=8, for 2014",
               note = "Data from SOM",
               output = "latex")
```


```{r echo=FALSE, include = TRUE}
table_proportions_cluster_2014k8
```


#### Descriptives of lifestyle and class variables

Here I create the descriptive tables which show the mean values of different lifestyle activities, as well as for class variables, for each cluster.

```{r}
# Choosing all variables, now adding class variables and demographics.

data2014_tabledata_k8 <- data2014_k8 %>% 
  select(clusters_2014_k8,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1,
         euroesec_2,
         Income,
         Educationlevel,
         Subjective_class_today,
         age,
         sex_binary)


```

```{r}
# Converting to data frame.

data2014_tabledata_k8 <- as.data.frame(data2014_tabledata_k8)

# Making variables numeric to calculate means.

data2014_tabledata_means_k8 <- data2014_tabledata_k8 %>%
  mutate_all(as.numeric)

# Making cluster names character.
data2014_tabledata_means_k8 <- data2014_tabledata_means_k8 %>%
  mutate(clusters_2014_k8 = as.character(clusters_2014_k8))

```

```{r echo=FALSE, include = TRUE}
# Table with descriptive statistics of the clusters using Mean. Output in latex for Overleaf.


datasummary(praying+
         Cinema+
         Rock_pop+
         Theatre+
         Classical_opera+
         Library+
         Pub_restaurant+
         Drawn_painted+
         Friends+
         Church+
         snus+
         gamble_lottery+
         p1+
         euroesec_2+
         Income+
         Educationlevel+
         Subjective_class_today+
         age+
         sex_binary ~ clusters_2014_k8 * (Mean) + Mean,
            data = data2014_tabledata_means_k8,
         title = "2014",
         output = "latex")



```
#### Descriptive table

```{r}
# Descriptives showing the percentages of observations that reside in each cluster. Output in latex for Overleaf.

datasummary(
         sex_binary ~ clusters_2014_k8 * (Percent()) + N, #I use sex_binary as a placeholder to create a table. 
            data = data2014_tabledata_means_k8,
         title = "2014",
         output = "latex")

```


## 2019


```{r}
# Cut dendrogram into 8 clusters representing the number of parties.

clusters_2019_k8 <- cutree(cluster_result_2019, k = 8)

clusters_2019_k8 <- as.data.frame(clusters_2019_k8)

# Setting rownames as idnr, thereby we can merge together with the original data on idnr.

clusters_2019_k8$idnr <- row.names(clusters_2019_k8)

row.names(clusters_2019_k8) <- NULL
```

```{r}
# Taking the original data, filtered on year 2014.

data2019_k8 <- data %>% 
  filter(year == 2019)
```

```{r}
# Merging cluster assignment with original data by idnr.

data2019_k8 <- merge(data2019_k8, clusters_2019_k8, by = "idnr")
```


#### Party proportions
```{r}
# Here I create separate vectors for each cluster to be able to calculate proportions.

Clusteringdata_2019k8_cluster1 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "1")

Clusteringdata_2019k8_cluster2 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "2")

Clusteringdata_2019k8_cluster3 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "3")

Clusteringdata_2019k8_cluster4 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "4")

Clusteringdata_2019k8_cluster5 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "5")

Clusteringdata_2019k8_cluster6 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "6")

Clusteringdata_2019k8_cluster7 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "7")

Clusteringdata_2019k8_cluster8 <- data2019_k8 %>% 
  filter(clusters_2019_k8 == "8")
```

```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc1_k8 <- prop.table(table(Clusteringdata_2019k8_cluster1$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc1_k8 <- as.data.frame(category_proportionsc1_k8)

names(category_proportions_dfc1_k8) <- c("Party", "Proportion_C1")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 2
category_proportionsc2_k8 <- prop.table(table(Clusteringdata_2019k8_cluster2$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc2_k8 <- as.data.frame(category_proportionsc2_k8)

names(category_proportions_dfc2_k8) <- c("Party", "Proportion_C2")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 3
category_proportionsc3_k8 <- prop.table(table(Clusteringdata_2019k8_cluster3$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc3_k8 <- as.data.frame(category_proportionsc3_k8)

names(category_proportions_dfc3_k8) <- c("Party", "Proportion_C3")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 4
category_proportionsc4_k8 <- prop.table(table(Clusteringdata_2019k8_cluster4$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc4_k8 <- as.data.frame(category_proportionsc4_k8)

names(category_proportions_dfc4_k8) <- c("Party", "Proportion_C4")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 5
category_proportionsc5_k8 <- prop.table(table(Clusteringdata_2019k8_cluster5$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc5_k8 <- as.data.frame(category_proportionsc5_k8)

names(category_proportions_dfc5_k8) <- c("Party", "Proportion_C5")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc6_k8 <- prop.table(table(Clusteringdata_2019k8_cluster6$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc6_k8 <- as.data.frame(category_proportionsc6_k8)

names(category_proportions_dfc6_k8) <- c("Party", "Proportion_C6")

```

```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc7_k8 <- prop.table(table(Clusteringdata_2019k8_cluster7$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc7_k8 <- as.data.frame(category_proportionsc7_k8)

names(category_proportions_dfc7_k8) <- c("Party", "Proportion_C7")

```


```{r}
# Calculating proportions of categories from cb10 for cluster 1
category_proportionsc8_k8 <- prop.table(table(Clusteringdata_2019k8_cluster8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfc8_k8 <- as.data.frame(category_proportionsc8_k8)

names(category_proportions_dfc8_k8) <- c("Party", "Proportion_C8")

```


```{r}
# Calculate proportions of categories from cb10 for all

category_proportionsk8 <- prop.table(table(data2019_k8$cb10))

# Converting the result to a data frame for plotting
category_proportions_dfk8 <- as.data.frame(category_proportionsk8)

names(category_proportions_dfk8) <- c("Party", "Proportion_all")
```

```{r}
# Merging all proportion data frames together.

merged_2019_k8 <- merge(category_proportions_dfc1_k8, category_proportions_dfc2_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc3_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc4_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc5_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc6_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc7_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfc8_k8, by = "Party")

merged_2019_k8 <- merge(merged_2019_k8 , category_proportions_dfk8, by = "Party")

```

```{r}
# Renaming party codes to full party names.

merged_2019_k8 <- merged_2019_k8 %>% 
  mutate(Party = case_when(
    Party == 1 ~ "Left Party",
    Party == 2 ~ "Social Democrats",
    Party == 3 ~ "Centre Party",
    Party == 4 ~ "Liberals",
    Party == 5 ~ "Moderate Party",
    Party == 6 ~ "Christian Democrats",
    Party == 7 ~ "Green Party",
    Party == 10 ~ "Sweden Democrats"
  ))

# Renaming for clarity.

merged_2019_k8 <- merged_2019_k8 %>% 
  rename("Cluster 1 " = Proportion_C1,
         "Cluster 2 " = Proportion_C2,
         "Cluster 3 " = Proportion_C3,
         "Cluster 4 " = Proportion_C4,
         "Cluster 5 " = Proportion_C5,
         "Cluster 6 " = Proportion_C6,
         "Cluster 7 " = Proportion_C7,
         "Cluster 8 " = Proportion_C8,
         "All " = Proportion_all)

# Taking the proportion * 100 to get percentages.

merged_2019_k8 <- merged_2019_k8 %>%
  mutate_at(vars(starts_with("Cluster")), funs(. * 100))

merged_2019_k8 <- merged_2019_k8 %>%
  mutate_at(vars(starts_with("All")), funs(. * 100))

```

```{r}
# Printing the table for proportions of respondents in each cluster that prefer each party. Output in latex for Overleaf.

table_proportions_cluster_2019k8 <- datasummary_df(merged_2019_k8, fmt = 0,
               title = "Preferred party of respondents in each cluster, k=8, for 2019",
               note = "Data from SOM",
               output = "latex")
```


```{r echo=FALSE, include = TRUE}
table_proportions_cluster_2019k8
```


#### Descriptives of lifestyle and class variables

Here I create the descriptive tables which show the mean values of different lifestyle activities, as well as for class variables, for each cluster.

```{r}
# Choosing all variables, now adding class variables and demographics.

data2019_tabledata_k8 <- data2019_k8 %>% 
  select(clusters_2019_k8,
         praying,
         Cinema,
         Rock_pop,
         Theatre,
         Classical_opera,
         Library,
         Pub_restaurant,
         Drawn_painted,
         Friends,
         Church,
         snus,
         gamble_lottery,
         p1,
         euroesec_2,
         Income,
         Educationlevel,
         Subjective_class_today,
         age,
         sex_binary)


```

```{r}
# Converting to data frame.

data2019_tabledata_k8 <- as.data.frame(data2019_tabledata_k8)

# Making variables numeric to calculate means.

data2019_tabledata_means_k8 <- data2019_tabledata_k8 %>%
  mutate_all(as.numeric)

# Making cluster names character.

data2019_tabledata_means_k8 <- data2019_tabledata_means_k8 %>%
  mutate(clusters_2019_k8 = as.character(clusters_2019_k8))

```

```{r echo=FALSE, include = TRUE}
# Table with descriptive statistics of the clusters using Mean. Output in latex for Overleaf.

datasummary_table_2019_k8 <- datasummary(
  praying+
         Cinema+
         Rock_pop+
         Theatre+
         Classical_opera+
         Library+
         Pub_restaurant+
         Drawn_painted+
         Friends+
         Church+
         snus+
         gamble_lottery+
         p1+
         euroesec_2+
         Income+
         Educationlevel+
         Subjective_class_today+
         age+
         sex_binary ~ clusters_2019_k8 * (Mean) + Mean,
            data = data2019_tabledata_means_k8,
         title = "2019", 
  output = "latex")



```

```{r}
datasummary_table_2019_k8
```

#### Descriptive table

```{r}
# Descriptives showing the percentages of observations that reside in each cluster. Output in latex for Overleaf.

datasummary(
         sex_binary ~ clusters_2019_k8 * (Percent()) + N, #I use sex_binary as a placeholder to create a table. 
            data = data2019_tabledata_means_k8,
         title = "2019",
         output = "latex")

```



# Silhouette



```{r}
# Calculating silhouette scores
silhouette_score_2019 <- silhouette(clusters_2019_k8[,1], dist_matrix_2019)

silhouette_score_2019_table <- mean(silhouette_score_2019[,3])

silhouette_score_2014 <- silhouette(clusters_2014_k8[,1], dist_matrix_2014)

silhouette_score_2014_table <- mean(silhouette_score_2014[,3])

silhouette_score_2008 <- silhouette(clusters_2008_k8[,1], dist_matrix_2008)

silhouette_score_2008_table <- mean(silhouette_score_2008[,3])
```

```{r}
# Storing silhouette score in dataframe
silhouette_scores_df <- data.frame(
  Year = c(2008, 2014, 2019),
  Silhouette_Score = c(silhouette_score_2008_table, silhouette_score_2014_table, silhouette_score_2019_table)
)
```

```{r}
# Creating table for latex
datasummary_df(silhouette_scores_df,
               output = "latex")
```

