---
title: "Imputation_0405"
author: "Mattias Bengtsson"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(nnet)
library(missForest)
library(randomForest)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r}
# Setting working directory
setwd("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM")
```

```{r}
#Reading data
data <- read.csv("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Super-R-SOM (1986-2021).csv", header = TRUE, quote = "") 
```



```{r}
#Cleaning column names
colnames(data) <- gsub("^X|\\.*$", "", colnames(data)) 
colnames(data) <- gsub("\\.", "", colnames(data))
```
```{r}
#Removing columns without data
data <- data[, colSums(!is.na(data)) > 0] 

```

```{r}
#Choosing years 2001 to 2021
data20012021 <- data |> 
  filter(year >= 2001 & year <= 2021) 
```


```{r}
#Removing columns without any data
data20012021 <- data20012021[, colSums(!is.na(data20012021)) > 0] 
```



```{r}
#Choosing variables to be included. Here i chose more variables than i was going to use in the analysis. This ensure a better imputation process, but also since i at this stage was unsure of which variables to include. 

data_selected <- data20012021 |> select(idnr, year, lmsit, cityrur, citizen, subclh, subclg, lan, edu3, occgr, euroesec, hinc5rel, age, sex, natgeo, mc10c, mc10k, md10, mb99b, mb99a, mb99d, mb99e, ma50c, ma70b, ma40h, ma40c,  ma40b,  ma30c, ma30l, lb10a, lb10b, ma10a, lc10a, lc20a, la10, jc10a, jb19, genzukin, growupf, growupm, growupp, eg10, edu2, cb10, cb20, bc10, bc20, ba10, lb10d, lb10e, ld10a,lc10a,lc20a, le10c, le10d, unionm, ga10g, ga10j, ga10q, ha10, ha200d, fc900b, jf10a, jh10e)
```



```{r}


# Convert all columns to numeric in order to remove the missing values that have values rather than "NA". 

data_selected <- data_selected %>%
  mutate(across(everything(), as.numeric))

# Define values to be replaced with NA
replacement_values <- c(95, 96, 97, 98, 99, 30, 998, 999, 9999)

# Replacing replacement_values with NA in all columns except for the specified ones
data_selected <- data_selected %>%
  mutate(across(-c(idnr, natgeo, age, year), ~ifelse(. %in% replacement_values, NA, .)))


```

```{r}
# Here i mutate all variables back to be as factor, except for year, age and idnr which are numeric. 
data_selected <- mutate_at(data_selected, 
                           vars(-year, -age, -idnr), 
                           as.factor)
```

```{r}
# Merging two variables measuring the same activity, "Exercised". 
data_selected <- data |> 
  mutate(merged_ma30 = coalesce(ma30c, ma30l))
```

```{r}
# I have chosen to impute for each year separately, thus i create separate datasets for each year. 
data_selected2001 <- data_selected |> 
  filter(year == 2001)


data_selected2002 <- data_selected |> 
  filter(year == 2002)


data_selected2003 <- data_selected |> 
  filter(year == 2003)


data_selected2004 <- data_selected |> 
  filter(year == 2004)


data_selected2005 <- data_selected |> 
  filter(year == 2005)

data_selected2006 <- data_selected |> 
  filter(year == 2006)


data_selected2007 <- data_selected |> 
  filter(year == 2007)


data_selected2008 <- data_selected |> 
  filter(year == 2008)

data_selected2009 <- data_selected |> 
  filter(year == 2009)

data_selected2010 <- data_selected |> 
  filter(year == 2010)

data_selected2011 <- data_selected |> 
  filter(year == 2011)

data_selected2012 <- data_selected |> 
  filter(year == 2012)

data_selected2013 <- data_selected |> 
  filter(year == 2013)

data_selected2014 <- data_selected |> 
  filter(year == 2014)

data_selected2015 <- data_selected |> 
  filter(year == 2015)

data_selected2016 <- data_selected |> 
  filter(year == 2016)

data_selected2017 <- data_selected |> 
  filter(year == 2017)

data_selected2018 <- data_selected |> 
  filter(year == 2018)

data_selected2019 <- data_selected |> 
  filter(year == 2019)

data_selected2020 <- data_selected |> 
  filter(year == 2020)
data_selected2021 <- data_selected |> 
  filter(year == 2021)
```

Down below i use the missforest to impute. The same procedure is done for each year. I chose to not do it in a loop since it takes a long time for the imputation process and each year. Thus doing it in separate chunks was more managable for my computer. 

The first line "imputed_yearx <- missForest(data_selectedyearx, verbose = TRUE)",applies the missForest algorithm on the dataset "data_selectedyearx". "Verbose" tells it to provide a detailed logging of the process. This enables me to make sure it is working since each imputation takes a long time. The second line extract the imputed dataframe from the result of the missForest function and assign it to imputed_dataframeyearX. 

```{r}
#For 2001
imputed_2001 <- missForest(data_selected2001, verbose = TRUE)
imputed_dataframe2001 <- imputed_2001$ximp
```


```{r}
# For 2002
imputed_2002 <- missForest(data_selected2002, verbose = TRUE)
imputed_dataframe2002 <- imputed_2002$ximp
```


```{r}
# For 2003
imputed_2003 <- missForest(data_selected2003, verbose = TRUE)
imputed_dataframe2003 <- imputed_2003$ximp
```


```{r}
# For 2004
imputed_2004 <- missForest(data_selected2004, verbose = TRUE)
imputed_dataframe2004 <- imputed_2004$ximp
```


```{r}
# For 2005
imputed_2005 <- missForest(data_selected2005, verbose = TRUE)
imputed_dataframe2005 <- imputed_2005$ximp
```


```{r}
#For 2006
imputed_2006 <- missForest(data_selected2006, verbose = TRUE)
imputed_dataframe2006 <- imputed_2006$ximp
```


```{r}
# For 2007
imputed_2007 <- missForest(data_selected2007, verbose = TRUE)
imputed_dataframe2007 <- imputed_2007$ximp
```


```{r}
# For 2008
imputed_2008 <- missForest(data_selected2008, verbose = TRUE)
imputed_dataframe2008 <- imputed_2008$ximp
```


```{r}
#For 2009

imputed_2009 <- missForest(data_selected2009, verbose = TRUE)
imputed_dataframe2009 <- imputed_2009$ximp
```


```{r}
# For 2010
imputed_2010 <- missForest(data_selected2010, verbose = TRUE)
imputed_dataframe2010 <- imputed_2010$ximp
```


```{r}
# For 2011
imputed_2011 <- missForest(data_selected2011, verbose = TRUE)
imputed_dataframe2011 <- imputed_2011$ximp


```


```{r}
# For 2012
imputed_2012 <- missForest(data_selected2012, verbose = TRUE)
imputed_dataframe2012 <- imputed_2012$ximp
```


```{r}
# For 2013

imputed_2013 <- missForest(data_selected2013, verbose = TRUE)
imputed_dataframe2013 <- imputed_2013$ximp


```


```{r}
# For 2014
imputed_2014 <- missForest(data_selected2014, verbose = TRUE)
imputed_dataframe2014 <- imputed_2014$ximp
```


```{r}
# For 2015

imputed_2015 <- missForest(data_selected2015, verbose = TRUE)
imputed_dataframe2015 <- imputed_2015$ximp
```


```{r}
# For 2016

imputed_2016 <- missForest(data_selected2016, verbose = TRUE)
imputed_dataframe2016 <- imputed_2016$ximp
```


```{r}
# For 2017

imputed_2017 <- missForest(data_selected2017, verbose = TRUE)
imputed_dataframe2017 <- imputed_2017$ximp
```


```{r}
# For 2018

imputed_2018 <- missForest(data_selected2018, verbose = TRUE)
imputed_dataframe2018 <- imputed_2018$ximp
```


```{r}
# For 2019
imputed_2019 <- missForest(data_selected2019, verbose = TRUE)
imputed_dataframe2019 <- imputed_2019$ximp
```


```{r}
# For 2020

imputed_2020 <- missForest(data_selected2020, verbose = TRUE)
imputed_dataframe2020 <- imputed_2020$ximp
```


```{r}
# For 2021
imputed_2021 <- missForest(data_selected2021, verbose = TRUE)
imputed_dataframe2021 <- imputed_2021$ximp

```

```{r}
# Here i Save all of the imputed datasets into a list
imputed_data_list2001_2021 <- list(
  imputed_dataframe2001,
  imputed_dataframe2002,
  imputed_dataframe2003,
  imputed_dataframe2004,
  imputed_dataframe2005,
  imputed_dataframe2006,
  imputed_dataframe2007,
  imputed_dataframe2008,
  imputed_dataframe2009,
  imputed_dataframe2010,
  imputed_dataframe2011,
  imputed_dataframe2012,
  imputed_dataframe2013,
  imputed_dataframe2014,
  imputed_dataframe2015,
  imputed_dataframe2016,
  imputed_dataframe2017,
  imputed_dataframe2018,
  imputed_dataframe2019,
  imputed_dataframe2020,
  imputed_dataframe2021
)
# Then i combining each dataframe from the list to one
combined_data01_21_incomeandcult <- bind_rows(imputed_data_list2001_2021)

# Saving the dataframe to CSV. This is then the dataset i work with for every analysis
write.csv(combined_data01_21_incomeandcult, file = "combined_imputed_data0121_incomeandcult.csv")
```