---
title: "Random Forest Models"
author: "Mattias Bengtsson"
date: "2024-04-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE)
```

Here predictions are made with random forest for a Party preference (CB10), and Socialist/not socialist (constructed from CB10). 

# Setup

```{r}
library(ggplot2)
library(tidyverse)
library(randomForest)
library(modelsummary)
library(performanceEstimation)
```


```{r}
#Setting the work directory
#setwd("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions")#liu dator
```
```{r}
setwd("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions") #Egen dator
```


```{r}
#Loading the imputed dataset
#imputed_data_0121 <- read.csv("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/combined_imputed_data0121_incomeandcult.csv")

imputed_data_0121 <- read.csv("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/combined_imputed_data0121_incomeandcult.csv")
```

In this analysis I aim to test different operationalizations of the concept "Class" to evaluate if they differ in their ability to predict preferred party.

```{r}
data <- mutate_at(imputed_data_0121,
                           vars(-year, -age, -idnr), 
                           as.factor)
```

```{r}

# Filter the data to exclude parties that is not in the national parliament
data <- data %>% 
  filter(cb10 != "11" & cb10 != "12")

data$cb10 <- droplevels(data$cb10)

```


```{r}
# Constructing the "Socialist or not" variable, in accordance with definition in the thesis.
# 1 = socialist, 0 = non-socialist
data <- data %>% 
  mutate(socialist = case_when(
    cb10 == "1" ~ "1",  # Left party
    cb10 == "2" ~ "1",  # Social Democrats
    cb10 == "7" ~ "1",  # Green party
    TRUE ~ "0"
  ))
```

```{r}
# Making sure it is a factor variable 
data$socialist <- as.factor(data$socialist)
```


## Analysing data per year

```{r}
filtered_data <- subset(data, year >= 2008 & year <= 2021) # Filtering out data per year, 2008 to 2021. 

dataperyear <- table(filtered_data$year)

dataperyear <- data.frame(dataperyear)
```


```{r}
# Creating a plot with the amount of observations for each year. 

dataplot <- ggplot(dataperyear, aes(x = Var1, y = Freq)) +
     geom_bar(stat = "identity", fill = "lightblue", width = 0.7) +  # Bar plot
     theme_minimal() +
     labs(x = "Year", y = "Sample size", title = "Sample size by year after imputation", 
          caption = "Source: Authors' imputation using data from Göteborgs Universitet (2023)")

print(dataplot)
```


# CB10 as outcome variable 

Here i use CB10 as outcome variable (preference for party X). These models thus aim to answer research question Are lifestyle indicators increasingly predictive of party preference in Sweden?

For each model, the first loop predicts which party the respondents prefer for each year, and saves the results for each year. 

In the next loop, i retrieve the OOB results from each year in a dataframe. This dataframe thus contains the OOB error rate for each year. 

After that all models with different variable selections have run, i merge the OOB results from each different model into one dataframe. This dataframe is then used to create tables, and plots that is analysed. The same process is done for the models with the outcome variable "Socialist party". 

###  Lifestyle 

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets (filtered by year).
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  model_life <- randomForest(cb10 ~ mc10c +  ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
 # Saving the model with the corresponding name. This name is called in the next loop to retrieve the oob
  assign(paste("model_life", year, sep = ""), model_life)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_life)
}
```

```{r}

# Creating an empty dataframe to store the results
oob_datalife <- data.frame(year = numeric(), OOBlife = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_life", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datalife <- rbind(oob_datalife, data.frame(year = year, OOBlife = current_oob))
}

# Print the compiled dataframe
print(oob_datalife)

```


### Euroesec

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  model_euroesec <- randomForest(cb10 ~ euroesec, data = current_data)
  
  # Saving the model with the corresponding name
  assign(paste("model_euroesec", year, sep = ""), model_euroesec)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_euroesec)
}

```

```{r}
# Creating an empty dataframe to store the results
oob_dataeuroesec <- data.frame(year = numeric(), OOBeuroesec = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Extract the model for the current year
  current_model <- get(paste("model_euroesec", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_dataeuroesec <- rbind(oob_dataeuroesec, data.frame(year = year, OOBeuroesec = current_oob))
}

# Print the compiled dataframe
print(oob_dataeuroesec)


```

### Sublch 
```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets_subclh <- list()

# Loop through the years from 2001 to 2021
for (year in 2001:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets_subclh[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  modelsubclh <- randomForest(cb10 ~ subclh, data = current_data)
  
  # Saving the model with the corresponding name
  assign(paste("modelsubclh", year, sep = ""), modelsubclh)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclh)
}

```


```{r}
# Creating an empty dataframe to store the results
oob_datasubclh <- data.frame(year = numeric(), OOBsubclh = numeric())

# Loop through the years from 2001 to 2021
for (year in 2001:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelsubclh", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datasubclh <- rbind(oob_datasubclh, data.frame(year = year, OOBsubclh = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclh)

```

### Sublcg
```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets_subclg <- list()

# Loop through the years from 2001 to 2021
for (year in 2001:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets_subclg[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  modelsubclg <- randomForest(cb10 ~ subclg, data = current_data)
  
  # Saving the model with the corresponding name
  assign(paste("modelsubclg", year, sep = ""), modelsubclg)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclg)
}

```


```{r}
# Creating an empty dataframe to store the results
oob_datasubclg <- data.frame(year = numeric(), OOBsubclg = numeric())

# Loop through the years from 2001 to 2021
for (year in 2001:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelsubclg", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datasubclg <- rbind(oob_datasubclg, data.frame(year = year, OOBsubclg = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclg)


```



### Hinc5rel - income 

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets_hinc5 <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets_hinc5[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  modelhinc5 <- randomForest(cb10 ~ hinc5rel, data = current_data)
  
  # Saving the model with the corresponding name
  assign(paste("modelhinc5", year, sep = ""), modelhinc5)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelhinc5)
}

```


```{r}
# Creating an empty dataframe to store the results
oob_datahinc5 <- data.frame(year = numeric(), OOBhinc5 = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelhinc5", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datahinc5 <- rbind(oob_datahinc5, data.frame(year = year, OOBhinc5 = current_oob))
}

# Print the compiled dataframe
print(oob_datahinc5)

```

### Class + lifestyle 

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filtering the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Storing the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Training the random forest model for the current year
  model_classlifestyle <- randomForest(cb10 ~ euroesec + subclg + subclh + hinc5rel + edu3 + mc10c + mb99b + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Saving the model with the corresponding name
  assign(paste("model_classlifestyle", year, sep = ""), model_classlifestyle)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_classlifestyle)
}
```

```{r}

# Creating an empty dataframe to store the results
oob_dataclasslifestyle <- data.frame(year = numeric(), OOBclasslifestyle = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extracting the model for the current year
  current_model <- get(paste("model_classlifestyle", year, sep = ""))
  
  # Extracting the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_dataclasslifestyle <- rbind(oob_dataclasslifestyle, data.frame(year = year, OOBclasslifestyle = current_oob))
}



```



### Full class model

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_fullclass <- randomForest(cb10 ~ euroesec + subclg + subclh + hinc5rel + edu3, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_fullclass", year, sep = ""), model_fullclass)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_fullclass)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datafullclass <- data.frame(year = numeric(), OOBfullclass = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_fullclass", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datafullclass <- rbind(oob_datafullclass, data.frame(year = year, OOBfullclass = current_oob))
}



```



## Merging

Here I merge all the dataframes with the OOB-errors together. That way i can produce a table with them all together, and plots. 

```{r}
# Lifestyle model merged with euroesec
merged_generations_life <- merge(oob_datalife, oob_dataeuroesec)

# Adding subjective home class
merged_generations_life <- merge(merged_generations_life, oob_datasubclh)


# Adding Subjective class today
merged_generations_life <- merge(merged_generations_life, oob_datasubclg)


# Adding Income class
merged_generations_life <- merge(merged_generations_life, oob_datahinc5)

# Adding Full class model
merged_generations_life <- merge(merged_generations_life, oob_datafullclass)

# Adding the class and lifestyle model
merged_generations_life <- merge(merged_generations_life, oob_dataclasslifestyle)

```

```{r}
merged_generations_life
```


```{r}
# Output which is used to produce table in Overleaf
datasummary_df(merged_generations_life,
               output = "latex")
```


### Plotting

This plot shows the OOB-error rate over time, for each model, with party preference as outcome. 
```{r}

randomforestplot <- ggplot(merged_generations_life, aes(x = year)) +
  geom_line(aes(y = OOBlife, color = "Lifestyle"), linewidth = 1.3) +
 # geom_line(aes(y = OOBeuroesec, color = "Euroesec"), linewidth = 1.3)+
#geom_line(aes(y = OOBsubclg, color = "Subjective class today"), linewidth = 1.3)+
  #geom_line(aes(y = OOBsubclh, color = "Subjective class home"), linewidth = 1.3)+
   # geom_line(aes(y = OOBhinc5, color = "Income class"), linewidth = 1.3)+
  geom_line(aes(y = OOBfullclass, color = "Full classmodel"), linewidth = 1.3)+
  geom_line(aes(y = OOBclasslifestyle, color = "Class and Lifestyle model"), linewidth = 1.3)+
 labs(x = "Year", y = "Out-of-Bag Error (OOB)", color = "Models",
                 caption = "Data Source: Authors' imputation using data from SOM Institute, 2022") +theme_minimal()+
   ggtitle("Predictive Perfomance of Random Forest models on preference for party, per year") +
  scale_y_continuous(limits = c(0.4, 0.8)) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2))
  #xlim(2007, 2021)
```
```{r}
# Saving the plot for overleaf

ggsave("randomforestplot_cb10_ny.pdf", randomforestplot, width = 8, height = 6, units = "in", dpi = 300)

```

# Same for Socialist outcome
Down below the same routine is performed, but with the constructed "Socialist" variable. 
###  Lifestyle 

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_life_sc <- randomForest(socialist ~ mc10c + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_life_sc", year, sep = ""), model_life_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_life_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datalife_sc <- data.frame(year = numeric(), OOBlife_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_life_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datalife_sc <- rbind(oob_datalife_sc, data.frame(year = year, OOBlife_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datalife_sc)

```


### Euroesec

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_euroesec_sc <- randomForest(socialist ~ euroesec, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_euroesec_sc", year, sep = ""), model_euroesec_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_euroesec_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataeuroesec_sc <- data.frame(year = numeric(), OOBeuroesec_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_euroesec_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_dataeuroesec_sc <- rbind(oob_dataeuroesec_sc, data.frame(year = year, OOBeuroesec_sc = current_oob))
}

# Print the compiled dataframe
print(oob_dataeuroesec)

```

### Sublch 
```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclh_sc <- randomForest(socialist ~ subclh, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclh_sc", year, sep = ""), modelsubclh_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclh_sc)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclh_sc <- data.frame(year = numeric(), OOBsubclh_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelsubclh_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datasubclh_sc <- rbind(oob_datasubclh_sc, data.frame(year = year, OOBsubclh_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclh_sc)

```

### Sublcg
```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclg_sc <- randomForest(socialist ~ subclg, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclg_sc", year, sep = ""), modelsubclg_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclg_sc)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclg_sc <- data.frame(year = numeric(), OOBsubclg_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelsubclg_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datasubclg_sc <- rbind(oob_datasubclg_sc, data.frame(year = year, OOBsubclg_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclg_sc)

```



### Hinc5rel - income 

```{r}
set.seed(123)

# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelhinc5_sc <- randomForest(socialist ~ hinc5rel , data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelhinc5_sc", year, sep = ""), modelhinc5_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
print(modelhinc5_sc)
}
```


```{r}
# Create an empty dataframe to store the results
oob_datahinc5_sc <- data.frame(year = numeric(), OOBhinc5_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelhinc5_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datahinc5_sc <- rbind(oob_datahinc5_sc, data.frame(year = year, OOBhinc5_sc = current_oob))
}
```

### Full class model

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_fullclass_sc <- randomForest(socialist ~ euroesec + subclg + subclh + hinc5rel + edu3, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_fullclass_sc", year, sep = ""), model_fullclass_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_fullclass_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datafullclass_sc <- data.frame(year = numeric(), OOBfullclass_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_fullclass_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_datafullclass_sc <- rbind(oob_datafullclass_sc, data.frame(year = year, OOBfullclass_sc = current_oob))
}



```

### Class + lifestyle 

```{r}
set.seed(123)
# Creating an empty list to store the filtered datasets
filtered_datasets <- list()

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Removing columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_classlifestyle_sc <- randomForest(socialist ~ euroesec + subclg + subclh + hinc5rel + edu3 + mc10c + mb99b + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_classlifestyle_sc", year, sep = ""), model_classlifestyle_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_classlifestyle_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataclasslifestyle_sc <- data.frame(year = numeric(), OOBclasslifestyle_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_classlifestyle_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Storing the year and OOB error to the dataframe
  oob_dataclasslifestyle_sc <- rbind(oob_dataclasslifestyle_sc, data.frame(year = year, OOBclasslifestyle_sc = current_oob))
}



```


## Merging
Here I merge all the dataframes with the OOB-errors together. That way i can produce a table with them all together, and plots. 

```{r}
# Lifestyle model merged with euroesec
merged_generations_life_sc <- merge(oob_datalife_sc, oob_dataeuroesec_sc)

# Adding Subjective class home
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datasubclh_sc)


# Adding Subjective class today
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datasubclg_sc)


# Adding  Income class
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datahinc5_sc)

# Adding full class model
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datafullclass_sc)

# Adding class and lifestyle model
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_dataclasslifestyle_sc)

```

```{r}
merged_generations_life_sc
```

```{r}
# This Latex output is used to produce table in overleaf. 
datasummary_df(merged_generations_life_sc,
               output = "latex")
```


### Plotting

Plotting the OOB-error rate trend over time, with "Socialist" as outcome variable. 
```{r}

randomforestplot_sc <- ggplot(merged_generations_life_sc, aes(x = year)) +
  geom_line(aes(y = OOBlife_sc, color = "Lifestyle model"), linewidth = 1.3) +
  geom_line(aes(y = OOBfullclass_sc, color = "Full class model"), linewidth = 1.3) +
  geom_line(aes(y = OOBclasslifestyle_sc, color = "Class and lifestyle model"), linewidth = 1.3)+
  labs(x = "Year", y = "Out-of-Bag Error (OOB)", color = "Models",
                 caption = "Data Source: Authors' imputation using data from Som Institute, 2022") +theme_minimal()+
  ggtitle("Predictive Perfomance of Random Forest models on preference for Socialist party, per year") +
  scale_y_continuous(limits = c(0.2, 0.6)) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2))
  #xlim(2007, 2021)
```

```{r}
randomforestplot_sc
```


```{r}
# Saving the plot for overleaf

ggsave("randomforestplot_sc.pdf", randomforestplot_sc, width = 8, height = 6, units = "in", dpi = 300)

```

