---
title: "Random Forest Models"
author: "Mattias Bengtsson"
date: "2024-04-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = FALSE)
```

Here predictions are made with random forest for all generations on CB10 and Socialist/not socialist. 

# Setup

```{r}
library(ggplot2)
library(tidyverse)
library(randomForest)
library(modelsummary)
```


```{r}
#setwd("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions")#liu dator
```
```{r}
setwd("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions") #Egen dator
```


```{r}
#imputed_data_0121 <- read.csv("C:/Users/matbe280/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/combined_imputed_data0121_incomeandcult.csv")

imputed_data_0121 <- read.csv("C:/Users/admin/OneDrive - Linköpings universitet/Datawrangling SOM/Predictions/combined_imputed_data0121_incomeandcult.csv")
```

In this analysis I aim to test different operationalizations of the concept "Class" to evaluate if they differ in their ability to predict preferred party.

```{r}
data <- mutate_at(imputed_data_0121,
                           vars(-year, -age, -idnr), 
                           as.factor)
```

```{r}

# Filter the data to exclude parties that is not in the national parliament
data <- data %>% 
  filter(cb10 != "11" & cb10 != "12")

data$cb10 <- droplevels(data$cb10)

```


```{r}
#Constructing the Socialist or Not variable, 1 = socialist party, 0 = Not socialist. 
data <- data %>% 
  mutate(socialist = case_when(
    cb10 == "1" ~ "1",  
    cb10 == "2" ~ "1",  
    cb10 == "7" ~ "1",
    TRUE ~ "0"
  ))
```

```{r}
data$socialist <- as.factor(data$socialist)

class(data$socialist)
```

```{r}
library(performanceEstimation)
```


```{r}

filtered_data <- subset(data, year >= 2008 & year <= 2021)

dataperyear <- table(filtered_data$year)

dataperyear <- data.frame(dataperyear)





dataplot <- ggplot(dataperyear, aes(x = Var1, y = Freq)) +
     geom_bar(stat = "identity", fill = "lightblue", width = 0.7) +  # Bar plot
     theme_minimal() +
     labs(x = "Year", y = "Sample size", title = "Sample size by year after imputation", 
          caption = "Source: Authors' imputation using data from Göteborgs Universitet (2023)")

print(dataplot)
```


# CB10 as outcome variable 
###  Lifestyle 

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_life <- randomForest(cb10 ~ mc10c +  ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_life", year, sep = ""), model_life)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_life)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datalife <- data.frame(year = numeric(), OOBlife = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_life", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datalife <- rbind(oob_datalife, data.frame(year = year, OOBlife = current_oob))
}

# Print the compiled dataframe
print(oob_datalife)

```

```{r}
importance(model_life2011)
```

```{r}
top_variable_data_life <- data.frame(year = numeric(), top_variable_life = character())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_life", year, sep = ""))
  
  # Extract variable importance for the current model
  current_variable_importance <- importance(current_model)
  
  # Find the top variable for the current year
  top_variable_life <- rownames(current_variable_importance)[which.max(current_variable_importance)]
  
  # Append the year and top variable to the dataframe
  top_variable_data_life <- rbind(top_variable_data_life, data.frame(year = year, top_variable_life = top_variable_life))
}

# Print the compiled dataframe
print(top_variable_data_life)


```
### Euroesec

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_euroesec <- randomForest(cb10 ~ euroesec, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_euroesec", year, sep = ""), model_euroesec)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_euroesec)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataeuroesec <- data.frame(year = numeric(), OOBeuroesec = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_euroesec", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_dataeuroesec <- rbind(oob_dataeuroesec, data.frame(year = year, OOBeuroesec = current_oob))
}

# Print the compiled dataframe
print(oob_dataeuroesec)

```

### Sublch 
```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets_subclh <- list()

# Loop through the years from 2008 to 2016
for (year in 2001:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclh <- randomForest(cb10 ~ subclh, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclh", year, sep = ""), modelsubclh)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclh)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclh <- data.frame(year = numeric(), OOBsubclh = numeric())

# Loop through the years from 2008 to 2021
for (year in 2001:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelsubclh", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datasubclh <- rbind(oob_datasubclh, data.frame(year = year, OOBsubclh = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclh)

```

### Sublcg
```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets_subclg <- list()

# Loop through the years from 2008 to 2016
for (year in 2001:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclg <- randomForest(cb10 ~ subclg, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclg", year, sep = ""), modelsubclg)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclg)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclg <- data.frame(year = numeric(), OOBsubclg = numeric())

# Loop through the years from 2008 to 2021
for (year in 2001:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelsubclg", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datasubclg <- rbind(oob_datasubclg, data.frame(year = year, OOBsubclg = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclg)

```



### Hinc5rel - income 

```{r}
set.seed(123)

# Create an empty list to store the filtered datasets
filtered_datasets_hinc5 <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelhinc5 <- randomForest(cb10 ~ hinc5rel , data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelhinc5", year, sep = ""), modelhinc5)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
print(modelhinc5)
}
```


```{r}
# Create an empty dataframe to store the results
oob_datahinc5 <- data.frame(year = numeric(), OOBhinc5 = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Extract the model for the current year
  current_model <- get(paste("modelhinc5", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datahinc5 <- rbind(oob_datahinc5, data.frame(year = year, OOBhinc5 = current_oob))
}
```

### Class + lifestyle 

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_classlifestyle <- randomForest(cb10 ~ euroesec + subclg + subclh + hinc5rel + edu3 + mc10c + mb99b + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_classlifestyle", year, sep = ""), model_classlifestyle)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_classlifestyle)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataclasslifestyle <- data.frame(year = numeric(), OOBclasslifestyle = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_classlifestyle", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_dataclasslifestyle <- rbind(oob_dataclasslifestyle, data.frame(year = year, OOBclasslifestyle = current_oob))
}



```



### Full class model

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_fullclass <- randomForest(cb10 ~ euroesec + subclg + subclh + hinc5rel + edu3, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_fullclass", year, sep = ""), model_fullclass)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_fullclass)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datafullclass <- data.frame(year = numeric(), OOBfullclass = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_fullclass", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datafullclass <- rbind(oob_datafullclass, data.frame(year = year, OOBfullclass = current_oob))
}



```



## Merging

```{r}
merged_generations_life <- merge(oob_datalife, oob_dataeuroesec)

# Subjective clas shome
merged_generations_life <- merge(merged_generations_life, oob_datasubclh)


#Subjective class today
merged_generations_life <- merge(merged_generations_life, oob_datasubclg)


# Income class
merged_generations_life <- merge(merged_generations_life, oob_datahinc5)

# Full class
merged_generations_life <- merge(merged_generations_life, oob_datafullclass)

merged_generations_life <- merge(merged_generations_life, oob_dataclasslifestyle)

```

```{r}
merged_generations_life
```


```{r}
datasummary_df(merged_generations_life,
               output = "latex")
```


### Plotting

```{r}
# Using full sample sizes
randomforestplot <- ggplot(merged_generations_life, aes(x = year)) +
  geom_line(aes(y = OOBlife, color = "Lifestyle"), linewidth = 1.3) +
 # geom_line(aes(y = OOBeuroesec, color = "Euroesec"), linewidth = 1.3)+
#geom_line(aes(y = OOBsubclg, color = "Subjective class today"), linewidth = 1.3)+
  #geom_line(aes(y = OOBsubclh, color = "Subjective class home"), linewidth = 1.3)+
   # geom_line(aes(y = OOBhinc5, color = "Income class"), linewidth = 1.3)+
  geom_line(aes(y = OOBfullclass, color = "Full classmodel"), linewidth = 1.3)+
  geom_line(aes(y = OOBclasslifestyle, color = "Class and Lifestyle model"), linewidth = 1.3)+
 labs(x = "Year", y = "Out-of-Bag Error (OOB)", color = "Models",
                 caption = "Data Source: Authors' imputation using data from SOM Institute, 2022") +theme_minimal()+
   ggtitle("Predictive Perfomance of Random Forest models on preference for party, per year") +
  scale_y_continuous(limits = c(0.4, 0.8)) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2))
  #xlim(2007, 2021)
```
```{r}
# Saving the plot for overleaf

ggsave("randomforestplot_cb10_ny.pdf", randomforestplot, width = 8, height = 6, units = "in", dpi = 300)

```

# Same for Socialist outcome



###  Lifestyle 

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_life_sc <- randomForest(socialist ~ mc10c + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_life_sc", year, sep = ""), model_life_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_life_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datalife_sc <- data.frame(year = numeric(), OOBlife_sc = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_life_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datalife_sc <- rbind(oob_datalife_sc, data.frame(year = year, OOBlife_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datalife_sc)

```


### Euroesec

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_euroesec_sc <- randomForest(socialist ~ euroesec, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_euroesec_sc", year, sep = ""), model_euroesec_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_euroesec_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataeuroesec_sc <- data.frame(year = numeric(), OOBeuroesec_sc = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_euroesec_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_dataeuroesec_sc <- rbind(oob_dataeuroesec_sc, data.frame(year = year, OOBeuroesec_sc = current_oob))
}

# Print the compiled dataframe
print(oob_dataeuroesec)

```

### Sublch 
```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclh_sc <- randomForest(socialist ~ subclh, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclh_sc", year, sep = ""), modelsubclh_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclh_sc)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclh_sc <- data.frame(year = numeric(), OOBsubclh_sc = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelsubclh_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datasubclh_sc <- rbind(oob_datasubclh_sc, data.frame(year = year, OOBsubclh_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclh_sc)

```

### Sublcg
```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelsubclg_sc <- randomForest(socialist ~ subclg, data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelsubclg_sc", year, sep = ""), modelsubclg_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(modelsubclg_sc)
}
```


```{r}

# Create an empty dataframe to store the results
oob_datasubclg_sc <- data.frame(year = numeric(), OOBsubclg_sc = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelsubclg_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datasubclg_sc <- rbind(oob_datasubclg_sc, data.frame(year = year, OOBsubclg_sc = current_oob))
}

# Print the compiled dataframe
print(oob_datasubclg_sc)

```



### Hinc5rel - income 

```{r}
set.seed(123)

# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  modelhinc5_sc <- randomForest(socialist ~ hinc5rel , data = current_data)
  
  # Save the model with the corresponding name
  assign(paste("modelhinc5_sc", year, sep = ""), modelhinc5_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
print(modelhinc5_sc)
}
```


```{r}
# Create an empty dataframe to store the results
oob_datahinc5_sc <- data.frame(year = numeric(), OOBhinc5_sc = numeric())

# Loop through the years from 2008 to 2021
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("modelhinc5_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datahinc5_sc <- rbind(oob_datahinc5_sc, data.frame(year = year, OOBhinc5_sc = current_oob))
}
```

### Full class model

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_fullclass_sc <- randomForest(socialist ~ euroesec + subclg + subclh + hinc5rel + edu3, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_fullclass_sc", year, sep = ""), model_fullclass_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_fullclass_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_datafullclass_sc <- data.frame(year = numeric(), OOBfullclass_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_fullclass_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_datafullclass_sc <- rbind(oob_datafullclass_sc, data.frame(year = year, OOBfullclass_sc = current_oob))
}



```

### Class + lifestyle 

```{r}
set.seed(123)
# Create an empty list to store the filtered datasets
filtered_datasets <- list()

# Loop through the years from 2008 to 2016
for (year in 2008:2021) {
  # Skip year 2016
  if (year == 2016) {
    next  # Skip the current iteration and move to the next year
  }
  
  # Filter the dataset for the current year
  current_data <- data[data$year == year, ]
  
  # Remove columns with no data
  current_data <- current_data[, colSums(!is.na(current_data)) > 0]
  
  # Store the filtered dataset in the list
  filtered_datasets[[as.character(year)]] <- current_data
  
  # Train the random forest model for the current year
  model_classlifestyle_sc <- randomForest(socialist ~ euroesec + subclg + subclh + hinc5rel + edu3 + mc10c + mb99b + ma50c + ma40h + lb10a + ma10a + la10 + jc10a + eg10 + bc10 + bc20 + lb10e + lb10b + le10c, data  = current_data)
  
  # Save the model with the corresponding name
  assign(paste("model_classlifestyle_sc", year, sep = ""), model_classlifestyle_sc)
  
  # Print the summary of the model
  print(paste("Random Forest Model for Year", year))
  print(model_classlifestyle_sc)
}
```

```{r}

# Create an empty dataframe to store the results
oob_dataclasslifestyle_sc <- data.frame(year = numeric(), OOBclasslifestyle_sc = numeric())

# Looping through the years from 2008 to 2021
for (year in 2008:2021) {
  if (year == 2016) {
    next  # Skip the current iteration if 2016 and move to the next year
  }
  # Extract the model for the current year
  current_model <- get(paste("model_classlifestyle_sc", year, sep = ""))
  
  # Extract the OOB error for the current model
  current_oob <- current_model$err.rate[nrow(current_model$err.rate), "OOB"]
  
  # Append the year and OOB error to the dataframe
  oob_dataclasslifestyle_sc <- rbind(oob_dataclasslifestyle_sc, data.frame(year = year, OOBclasslifestyle_sc = current_oob))
}



```


## Merging

```{r}
merged_generations_life_sc <- merge(oob_datalife_sc, oob_dataeuroesec_sc)

# Subjective clas shome
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datasubclh_sc)


#Subjective class today
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datasubclg_sc)


# Income class
merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datahinc5_sc)


merged_generations_life_sc <- merge(merged_generations_life_sc, oob_datafullclass_sc)

merged_generations_life_sc <- merge(merged_generations_life_sc, oob_dataclasslifestyle_sc)

```

```{r}
merged_generations_life_sc
```

```{r}
datasummary_df(merged_generations_life_sc,
               output = "latex")
```


### Plotting

```{r}
# Using full sample sizes
randomforestplot_sc <- ggplot(merged_generations_life_sc, aes(x = year)) +
  geom_line(aes(y = OOBlife_sc, color = "Lifestyle model"), linewidth = 1.3) +
  geom_line(aes(y = OOBfullclass_sc, color = "Full class model"), linewidth = 1.3) +
  geom_line(aes(y = OOBclasslifestyle_sc, color = "Class and lifestyle model"), linewidth = 1.3)+
  labs(x = "Year", y = "Out-of-Bag Error (OOB)", color = "Models",
                 caption = "Data Source: Authors' imputation using data from Som Institute, 2022") +theme_minimal()+
  ggtitle("Predictive Perfomance of Random Forest models on preference for Socialist party, per year") +
  scale_y_continuous(limits = c(0.2, 0.6)) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2))
  #xlim(2007, 2021)
```

```{r}
randomforestplot_sc
```


```{r}
# Saving the plot for overleaf

ggsave("randomforestplot_sc.pdf", randomforestplot_sc, width = 8, height = 6, units = "in", dpi = 300)

```

